### 2008-03-21 trunk/Papers/dl-gre-08/introduction.tex
intro

The generation of referring expressions (GRE) is one of the most
active and successful research areas in natural language generation.
Starting with Dale and Reiter's incremental algorithm \cite{Dale1995},
various researchers have added extensions such as reference to sets
\cite{Stone2000}, more expressive logical connectives
\cite{deemter01:_gener_refer_expres}, and relational expressions
\cite{dale91:_gener_refer_expres_invol_relat}.

@@

In this paper, we propose a fundamentally new way of looking at the
problem of generating referring expressions: We take a referring
expression describing an individual $i$ to be a formula of some
description logic (DL) which is satisfied by $i$ and no other
individual in the universe.  This allows us to organize existing
approaches to GRE in an expressiveness hierarchy.  For instance, the
classical Dale \& Reiter algorithms compute concepts in a description
logic that consists only of conjunctions of atoms, whereas Dale \&
Haddock and other relational algorithms compute concepts in the logic
\el, which permits conjunction and existential quantification.

Furthermore, we show how to addapt algorithms for computing the
\emph{bisimulation classes} of a DL model to the GRE problem.  Intuitively,
bisimulation classes are sets of individuals that satisfy exactly the
same concepts; if an individual is alone in its bisimulation class,
then the concept corresponding to this class represents a uniquely
referring expression.  We present bisimulation classes algorithms for
the description logics \el\ (only conjunction and existential
quantification) and \alc\ (full modal logic).  By essentially
computing referring expressions for all individuals in the domain at
once, these algorithms systematically avoid the infinite regress
problem.  The \el\ algorithm, which generates the same kind of
descriptions as Dale \& Haddock, is capable of generating 67\% of the
relational REs in the
\newcite{viethen06:_algor_for_gener_refer_expres} dataset correctly,
and takes about 400 ms to compute relational REs for all 100
individuals in a random scene.

The paper is structured as follows.  In Section~\ref{sec:bisim}, we
will first define description logic and fundamental concepts such as
bisimulation.  We will then show how to compute bisimulation classes
 and how to apply these algorithms to GRE in Section~\ref{sec:gre}.  In
Section~\ref{sec:discussion}, we evaluate our algorithms and discuss
our results.  Section~\ref{sec:related} compares our approach to
related research; in particular, it shows how various prominent GRE
algorithms can be seen as special cases of ours.
Section~\ref{sec:conclusion} concludes and points to future work.

>>>>

The generation of referring expressions (GRE) is one of the most
active and successful research areas in natural language generation.
Building upon Dale and Reiter's work \cite{dale89cooking,Dale1995},
various researchers have added extensions such as reference to sets
\cite{Stone2000}, more expressive logical connectives
\cite{deemter01:_gener_refer_expres}, and relational expressions
\cite{dale91:_gener_refer_expres_invol_relat}.

@@

In this paper, we propose to view the problem of generating referring
expressions (REs) as the problem of computing a formula of description
logic (DL) that denotes exactly the set of individuals that we want to
refer to.  This very natural idea has been mentioned in passing before
\cite{Krahmer2003,gardent07:_gener_bridg_defin_descr}; however, we
take it one step further by proposing DL as an interlingua for
comparing the REs produced by different approaches to GRE.  In this
way, we can organize existing GRE approaches in an expressiveness
hierarchy.  For instance, the classical Dale and Reiter algorithms
compute purely conjunctive formulas;
\newcite{deemter01:_gener_refer_expres} extends this language by
adding the other propositional connectives, whereas
\newcite{dale91:_gener_refer_expres_invol_relat} extends it by
allowing existential quantification.

Furthermore, the view of GRE as a problem of computing DL formulas
with a given extension allows us to apply existing algorithms for the
latter problem to obtain efficient algorithms for GRE.  We present
algorithms that compute such formulas for the description logics \el\
(which allows only conjunction and existential quantification) and
\alc\ (which also allows negation).  These algorithms effectively
compute REs for all individuals in the domain at the same time, which
allows them to systematically avoid the infinite regress problem.  The
\el\ algorithm is capable of generating 67\% of the relational REs in
the \newcite{viethen06:_algor_for_gener_refer_expres} dataset, in about
15 milliseconds.  The \alc\ algorithm is even faster; it computes
relational REs for all 100 individuals in a random model in 140
milliseconds.

The paper is structured as follows.  In Section~\ref{sec:bisim}, we
will first define description logics.  We will then show how to
generate REs by computing DL similarity sets for \alc\ and \el\ in
Section~\ref{sec:gre}.  In Section~\ref{sec:discussion}, we evaluate
our algorithms and discuss our results.  Section~\ref{sec:related}
compares our approach to related research; in particular, it shows how
various prominent GRE algorithms fit into the DL framework.
Section~\ref{sec:conclusion} concludes and points to future work.
